---
title: "ROAR-SRE-WJ-Analysis"
author: "Jasmine Tran"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

# Load Packages

Let's first load the packages that we need for this chapter.

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library(knitr)
library(stringr)
library(ggpubr)
library(mirt)
library(kableExtra)
library(tidyverse)
library(purrr)
library(plyr)
library(ggplot2)
library(viridis)
library(ggsci)
library(wesanderson)
library(dplyr)
library(plotrix)
opts_knit$set(root.dir = "~/Documents/Code/ROAR_organizer/ROAR-Score/")
```

## Settings
```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")
# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

## Functions
```{r}
correlate_rows <- function(x, y) {
  cor(x, y)
}
```

# Data 

## Data Loading (jsPsych Raw Trials)
```{r}
df.raw <- read.csv("~/Documents/roar-analysis/data/lmb_F1_C3_sre_data.csv")
df.wj.raw <- read.csv("~/Documents/roar-analysis/data/lmb_F1_C3_wj_srf_data.csv")
```

## Data Cleaning & Wrangling
```{r}
df.cleaned <- df.raw %>% 
  filter(studyId == "school-default", 
         task == "test_response", 
         completed == "True") %>% 
  mutate(correct = as.numeric(correct),
         gradeResponse = str_extract(userInfo, "(?<=grade': ').*?(?='\\})"))
```

```{r}
df.wj.cleaned <- df.wj.raw %>%
  select(-redcapId) %>%
  mutate(srfScore180 = srfCorrect - srfIncorrect) 
```

```{r}
df.not.guessing <- df.cleaned %>%
  group_by(pid, runId) %>%
  dplyr::summarise(medianRT = median(rt)) %>%
  filter(medianRT > 500) %>%
  mutate(keep = 1)
```

```{r}
df.keep <- df.cleaned %>%
  left_join(df.not.guessing, by=c("pid", "runId")) %>%
  filter(keep == 1) %>%
  select(-keep)
```

```{r}
df.ai.keep <- df.keep %>%
  filter(blockId == "ai",
         trialNumBlock == 1) %>%
  mutate(keep = 1) %>%
  select(pid, runId, gradeResponse, keep)
```

```{r}
df.keep.v3 <- df.keep %>%
  left_join(df.ai.keep, by=c("pid", "gradeResponse", "runId")) %>%
  filter(keep == 1)
```

```{r}
# this analysis has no tosrec data but leaving this just in case
#df.analysis <- rbind(df.keep.v2, df.keep.v3)

df.analysis <- df.keep.v3
```

```{r}
df.init.time <- df.analysis %>% 
  filter(trialNumBlock == 1) %>% 
  select(pid, runId, corpusId, time_elapsed) %>% 
  dplyr :: rename(init_time = time_elapsed) 
```

```{r}
df.block.time <- df.analysis %>% 
  select(pid, runId, gradeResponse, corpusId, blockId, itemId, correct, rt,  time_elapsed) %>% 
  left_join(df.init.time, by = c("pid", "runId", "corpusId")) %>% 
  mutate(minus_time = time_elapsed - init_time) %>% 
  mutate(part = as.numeric(case_when(
    minus_time <= 90000 ~ 90,
    minus_time > 90000 & minus_time < 120000 ~ 120,
    minus_time >= 120000 ~ 180))) %>% # Default case if none of the above conditions are met
  filter(!is.na(pid), pid != "") %>%
  mutate(correct_new = ifelse(correct == 0, -1, 1))
```

## Creating Reference Data Frame to Set Intervals for Block Timing 
```{r}
list <- seq(0, 180000, by = 10000)

intervals.ref <- data.frame(tibble(interval = list))
full.int.df <- df.block.time %>%
  dplyr::rename(seconds_elapsed = minus_time) %>%
  group_by(pid, runId, blockId, seconds_elapsed) %>%
  tally() %>% 
  select(-n) %>%
  merge(intervals.ref, all = TRUE)
```

## Creating a Grade Data Frame to Reference Later On
```{r}
df.grade <- df.block.time %>%
  select(pid, runId, gradeResponse) %>%
  unique()
```

## Augmenting Data for Lab Block into 10 Second Intervals
```{r}
df.lab.int <- df.block.time %>%
  select(pid, runId, gradeResponse, blockId, correct, correct_new, minus_time) %>%
  filter(blockId == "lab") %>%
  group_by(pid, runId, blockId) %>%
  dplyr::arrange(minus_time, .by_group = T) %>%
  reframe(seconds_elapsed = minus_time,
         labScore = cumsum(correct_new))
```

```{r}
df.lab.int.v1 <- df.lab.int %>%
  right_join(full.int.df, by=c("pid", "runId", "blockId", "seconds_elapsed")) %>%
  mutate(diff = interval - seconds_elapsed) %>%
  filter(diff >= 0, blockId == "lab") %>%
  group_by(pid, runId, interval) %>% 
  slice_min(diff) %>%
  dplyr::rename("labSecondsElapsed" = seconds_elapsed) %>%
  ungroup()
```

```{r}
df.lab.180.score <- df.lab.int.v1 %>%
  filter(interval == 180000) %>%
  select(pid, runId, "labScore180" = labScore)
```

```{r}
df.lab.int.final <- df.lab.int.v1 %>%
  left_join(df.lab.180.score, by=c("pid", "runId")) %>%
  left_join(df.grade, by=c("pid", "runId")) %>%
  dplyr::rename("grade" = gradeResponse)
```

```{r}
df.lab.int.final
```

## Augmenting Data for AI Block into 10 Second Intervals
```{r}
df.ai.int <- df.block.time %>%
  select(pid, runId, gradeResponse, blockId, correct, correct_new, minus_time) %>%
  filter(blockId == "ai") %>%
  group_by(pid, runId, blockId) %>%
  dplyr::arrange(minus_time, .by_group = T) %>%
  reframe(seconds_elapsed = minus_time,
         aiScore = cumsum(correct_new))
```

```{r}
df.ai.int.v1 <- df.ai.int %>%
  right_join(full.int.df, by=c("pid", "runId", "blockId", "seconds_elapsed")) %>%
  mutate(diff = interval - seconds_elapsed) %>%
  filter(diff >= 0, blockId == "ai") %>%
  group_by(pid, runId, interval) %>% 
  slice_min(diff) %>%
  dplyr::rename("aiSecondsElapsed" = seconds_elapsed) %>%
  ungroup()
```

```{r}
df.ai.180.score <- df.ai.int.v1 %>%
  filter(interval == 180000) %>%
  select(pid, runId, "aiScore180" = aiScore)
```

```{r}
df.ai.int.final <- df.ai.int.v1 %>%
  left_join(df.ai.180.score, by=c("pid", "runId")) %>%
  left_join(df.grade, by=c("pid", "runId")) %>%
  dplyr::rename("grade" = gradeResponse)
```

```{r}
df.ai.int.final
```

## Making Data Frame for 3 minute Lab Score & 3 minute SRF Score correlation
```{r}
df.lab.srf.cor <- df.lab.int.final %>%
  select(pid, runId, labScore, labScore180,
         labSecondsElapsed, interval) %>%
  left_join(df.wj.cleaned, by = c("pid")) %>%
  filter(!is.na(srfScore180)) %>%
  select(pid, runId, labScore180, srfScore180) %>%
  unique()
```

## Combining Lab & WJ SRF Data Frame for Analysis
```{r}
df.lab.srf.int <- df.lab.int.final %>%
  select(pid, runId, labScore, labScore180,
         labSecondsElapsed, interval) %>%
  left_join(df.wj.cleaned, by = c("pid")) %>%
  filter(!is.na(srfScore180)) %>%
  dplyr::group_by(interval) %>%
  reframe(oneThreeCorr = cor(srfScore180, labScore),
          n = n(), 
          standardError = sqrt((1 - oneThreeCorr^2) / (n - 2)),
          errorBarsMax = oneThreeCorr + standardError,
          errorBarsMin = oneThreeCorr - standardError) %>% 
  mutate(errorBarsMaxAdj = ifelse(errorBarsMax > 1.00, 1.00, errorBarsMax))
```

## Combining Lab & AI Data Frame for Analysis
```{r}
df.ai.srf.int <- df.ai.int.final %>%
  select(pid, runId, aiScore, aiScore180,
         aiSecondsElapsed, interval) %>%
  left_join(df.wj.cleaned, by = c("pid")) %>%
  filter(!is.na(srfScore180)) %>%
  group_by(interval) %>%
  reframe(oneThreeCorr = cor(srfScore180, aiScore),
          n = n(), 
          standardError = sqrt((1 - oneThreeCorr^2) / (n - 2)),
          errorBarsMax = oneThreeCorr + standardError,
          errorBarsMin = oneThreeCorr - standardError) %>% 
  mutate(errorBarsMaxAdj = ifelse(errorBarsMax > 1.00, 1.00, errorBarsMax))
```

## Making Data Frame for 3 minute AI Score & 3 minute SRF Score correlation
```{r}
df.ai.srf.cor <- df.ai.int.final %>%
  select(pid, runId, aiScore, aiScore180,
         aiSecondsElapsed, interval) %>%
  left_join(df.wj.cleaned, by = c("pid")) %>%
  filter(!is.na(srfScore180)) %>%
  select(pid, runId, aiScore180, srfScore180) %>%
  unique()
```

## Combining Lab & AI Data Frame Along with Grade Response for Analysis
```{r}
df.lab.srf.grade <- df.lab.int.final %>%
  merge(df.ai.int.final, by=c("pid", "runId", "interval", "grade")) %>%
  select(pid, runId, grade, labScore, labScore180, aiScore, aiScore180,
         labSecondsElapsed, aiSecondsElapsed, interval) %>%
  group_by(interval, grade) %>%
  reframe(equalCorr = cor(labScore, aiScore),
          oneThreeCorr = cor(labScore180, aiScore),
          n = n(), 
          standardError = sqrt((1 - oneThreeCorr^2) / (n - 2)),
          errorBarsMax = oneThreeCorr + standardError,
          errorBarsMin = oneThreeCorr - standardError)
```

# Data Visualizations

## Histogram of Grade Range for Lab-TOSREC-AI Dataset
```{r}
ggplot(df.lab.tosrec.demo, aes(x = grade, y = n, fill = grade)) +
  geom_bar(stat = "identity", width = 1, colour="gray25") +
  geom_text(aes(label = n), vjust = -0.5, color = "black", size = 4, family = "Avenir") +
  xlab("Grade") +
  ylab("Count") +
  scale_y_continuous(limits = c(0, 400), breaks=seq(0,400,50), expand = c(0, 0)) +
  scale_fill_manual(name = "Grade", values = wes_palette("Zissou1", n=8, type = "continuous")) +
  ggtitle("Grade Distribution of Lab-TOSREC Dataset") +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12))

ggplot(df.lab.ai.demo, aes(x = grade, y = n, fill = grade)) +
  geom_bar(stat = "identity", width = 1, colour="gray25") +
  geom_text(aes(label = n), vjust = -0.5, color = "black", size = 4, family = "Avenir") +
  xlab("Grade") +
  ylab("Count") +
  scale_y_continuous(limits = c(0, 85), breaks=seq(0,85,10), expand = c(0, 0)) +
  scale_fill_manual(name = "Grade", values = wes_palette("Zissou1", n=8, type = "continuous")) +
  ggtitle("Grade Distribution of Lab-AI Dataset") +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12))
```

## Inspecting Lab-SRF-AI correlations against block time in 10 second intervals
```{r}
ggplot(df.lab.srf.int, mapping = aes(x = interval,
                     y = oneThreeCorr)) + 
  geom_point(color="gray21") + 
  labs(title = "Lab-WJ SRF Correlation vs. Time Elasped", 
       x = "Time Elasped (in seconds)",
       y = "SRE - WJ SRF (3 minutes) Correlation") +
  geom_errorbar(aes(ymin = errorBarsMin, ymax = errorBarsMaxAdj), width = 0.5) +
  geom_smooth(method = "loess", color="darkseagreen", se=F) +
  scale_x_continuous(limits=c(0,180000), breaks=seq(0,180000,10000), labels=seq(0, 180, by=10)) +
  scale_y_continuous(limits = c(-0.4, 1), breaks=seq(-0.4,1.0,0.1)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12))

ggplot(df.ai.srf.int, mapping = aes(x = interval,
                     y = oneThreeCorr)) + 
  geom_point(color="gray21") + 
  labs(title = "AI-WJ SRF Correlation vs. Time Elasped", 
       x = "Time Elasped (in seconds)",
       y = "WJ SRF (3 minutes) - AI Correlation") +
  geom_errorbar(aes(ymin = errorBarsMin, ymax = errorBarsMaxAdj), width = 0.5) +
  geom_smooth(method = "loess", color="darkseagreen", se=F) +
  scale_x_continuous(limits=c(0,180000), breaks=seq(0,180000,10000), labels=seq(0, 180, by=10)) +
  scale_y_continuous(limits = c(0, 1), breaks=seq(0,1.0,0.1)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12))
```

## Inspecting Lab-TOSREC-AI correlations against block time in 10 second intervals (with Grade Response)
```{r}
ggplot(df.lab.tosrec.grade, mapping = aes(x = interval,
                     y = oneThreeCorr, color=grade)) + 
  labs(title = "Lab-TOSREC Correlation vs. Time Elasped", 
       x = "Time Elasped (in seconds)",
       y = "SRE - TOSREC (3 minutes) Correlation") +
  facet_wrap(~grade) +
  scale_color_manual(name = "Grade", values = wes_palette("Zissou1", n = 8, type = "continuous")) +
  geom_errorbar(aes(ymin = errorBarsMin, ymax = errorBarsMax), width = 0.5) +
  geom_smooth(method = "loess", se=F) +
  scale_x_continuous(limits=c(0,180000), breaks=seq(0,180000,30000), labels=seq(0, 180, by=30)) +
  scale_y_continuous(limits = c(0, 1), breaks=seq(0,1.0,0.2)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12)) 

ggplot(df.lab.ai.grade, mapping = aes(x = interval,
                     y = oneThreeCorr, color=grade)) + 
  labs(title = "Lab-AI Correlation vs. Time Elasped", 
       x = "Time Elasped (in seconds)",
       y = "Lab (3 minutes) - AI Correlation") +
  facet_wrap(~grade) +
  scale_color_manual(name = "Grade", values = wes_palette("Zissou1", n = 8, type = "continuous")) +
  geom_errorbar(aes(ymin = errorBarsMin, ymax = errorBarsMax), width = 0.5) +
  geom_smooth(method = "loess", se=F) +
  scale_x_continuous(limits=c(0,180000), breaks=seq(0,180000,30000), labels=seq(0, 180, by=30)) +
  scale_y_continuous(limits = c(0, 1), breaks=seq(0,1.0,0.2)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12)) 
```

## Inspecting full 3 minute Lab-AI-SRF correlations 
```{r}
ggplot(df.lab.srf.cor, mapping = aes(x = labScore180,
                     y = srfScore180)) + 
  geom_point() + 
  labs(title = "SRE lab score (3 minutes) vs. WJ SRF score (3 minutes)", 
       x = "Lab score (correct - incorrect)",
       y = "WJ SRF score (correct - incorrect)") +
  #scale_color_viridis(option = 'viridis') +
  geom_smooth(method = "lm", color="black") +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..), color = "black", 
           geom = "label", hjust = -0.3, vjust = 1) 

ggplot(df.ai.srf.cor, mapping = aes(x = aiScore180,
                     y = srfScore180)) + 
  geom_point() + 
  labs(title = "SRE AI score (3 minutes) vs. WJ SRF score (3 minutes)", 
       x = "AI score (correct - incorrect)",
       y = "WJ SRF score (correct - incorrect)") +
  #scale_color_viridis(option = 'viridis') +
  geom_smooth(method = "lm", color="black") +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..), color = "black", 
           geom = "label", hjust = -0.3, vjust = 1) 
```

```{r}
# df.combined.interval.grade <- df.lab.interval %>%
#   merge(df.tosrec.interval, by=c("pid", "runId", "interval")) %>%
#   select(pid, runId, labScore, tosrecScore, 
#          seconds_elasped_lab, seconds_elasped_tosrec, interval) %>%
#   left_join(df.grade, by=c("pid", "runId")) %>%
#   group_by(interval, gradeResponse) %>%
#   reframe(corr = cor(labScore, tosrecScore))
```

```{r}
# df.grade <- df.block.time %>%
#   select(pid, runId, gradeResponse)
```
