---
title: "ROAR-SRE-WJ-Analysis"
author: "Jasmine Tran"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

# Load Packages

Let's first load the packages that we need for this chapter.

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library(knitr)
library(stringr)
library(ggpubr)
library(mirt)
library(kableExtra)
library(tidyverse)
library(purrr)
library(plyr)
library(ggplot2)
library(viridis)
library(ggsci)
library(wesanderson)
library(dplyr)
library(plotrix)
library(reshape2)
library(gridExtra)
opts_knit$set(root.dir = "~/Documents/roar-analysis")
```

## Settings
```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")
# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

## Functions 
```{r}
# function to get upper triangle (?)
get_upper_tri = function(cormat) {
  cormat[lower.tri(cormat)] = NA
  return(cormat) }
```

# Data 

## Data location: https://drive.google.com/drive/u/1/folders/0AAol-yxLe1sfUk9PVA

## Data Loading (LMB participants)
```{r}
# data from study 1 (LMB participants)
df.lmb.sre.raw <- read.csv("~/Documents/roar-analysis/data/lmb_F1_C3_sre_data.csv")
df.lmb.wj.raw <- read.csv("~/Documents/roar-analysis/data/lmb_F1_C3_wj_srf_data.csv") 
```

## Data Loading (BRSSD participants)
```{r}
# data from study 2 (brssd validation)
df.brs.sre.raw <- read.csv("~/Documents/roar-analysis/data/brs_sre_data_trials_fall_23.csv")
df.brs.wj.raw <- read.csv("~/Documents/roar-analysis/data/brs_fall_23_validation_data.csv")
```

## Data Cleaning & Wrangling

### Processing raw SRE trials from JsPsych
```{r}
# filtering for completed runs and the correct participants
df.lmb.sre.cleaned <- df.lmb.sre.raw %>% 
  filter(grepl("rc-", pid), 
         task == "test_response", 
         completed == "True") %>% 
  mutate(correct = as.numeric(correct),
         gradeResponse = str_extract(userInfo, "(?<=grade': ').*?(?='\\})"))
```

```{r}
# filtering for completed runs and the correct participants
df.brs.sre.cleaned <- df.brs.sre.raw %>% 
  select(-pid) %>%
  dplyr::rename("pid" = user.assessmentPid) %>%
  filter(grepl("brs-", pid), 
         completed == "True",
         corpusId != "practice")
```

```{r}
# determining which participants are guessing with low accuracy and very fast response times
df.lmb.sre.guessing <- df.lmb.sre.cleaned %>%
  group_by(pid, runId, blockId) %>%
  dplyr::summarise(medianRT = median(rt),
                   correct = sum(correct),
                   attempted = n(), 
                   incorrect = attempted - correct, 
                   pCorr = correct/attempted) %>%
  filter((medianRT <= 1000 | medianRT >= 20000) & pCorr <= 0.65) %>%
  select(pid, runId) %>%
  mutate(remove = 1) %>%
  unique() %>%
  ungroup() 
```

```{r}
# determining which participants are guessing with low accuracy and very fast response times
df.brs.sre.guessing <- df.brs.sre.cleaned %>%
  group_by(pid, runId, corpusId) %>%
  dplyr::summarise(medianRT = median(rt),
                   correct = sum(correct),
                   attempted = n(), 
                   incorrect = attempted - correct, 
                   pCorr = correct/attempted) %>%
  filter((medianRT <= 1000 | medianRT >= 20000) & pCorr <= 0.65) %>%
  select(pid, runId) %>%
  mutate(remove = 1) %>%
  unique() %>%
  ungroup()
```

```{r}
# filtering out guessing participants
df.lmb.sre.runs <- df.lmb.sre.cleaned %>%
  left_join(df.lmb.sre.guessing, by=c("pid", "runId")) %>%
  filter(is.na(remove)) %>%
  select(-remove) %>%
  mutate(source = "LMB") %>%
  select(pid, runId, blockId, corpusId, gradeResponse,
         trialNumBlock, trial_index, time_elapsed,
         item, rt, correct, source)
```

```{r}
# filtering out guessing participants
df.brs.sre.runs <- df.brs.sre.cleaned %>%
  left_join(df.brs.sre.guessing, by=c("pid", "runId")) %>%
  filter(is.na(remove)) %>%
  select(-remove) %>%
  mutate(source = "BRSSD") %>%
  select(pid, runId, "blockId" = subtask, corpusId, 
         "gradeResponse" = user.grade,
         trialNumBlock, trial_index, time_elapsed,
         item, rt, correct, source)
```

### Merging the SRE data for LMB and BRSSD participants
```{r}
df.combined.sre.runs <- rbind(df.brs.sre.runs, df.lmb.sre.runs)
```

## Processing combined dataframe for timing analysis
```{r}
df.init.time <- df.combined.sre.runs %>% 
  filter(trialNumBlock == 1) %>% 
  select(pid, runId, corpusId, time_elapsed) %>% 
  dplyr :: rename(init_time = time_elapsed) 
```

```{r}
df.block.time <- df.combined.sre.runs %>% 
  select(pid, runId, gradeResponse, corpusId, blockId, item, correct, rt, time_elapsed, source) %>% 
  left_join(df.init.time, by = c("pid", "runId", "corpusId")) %>% 
  mutate(minus_time = time_elapsed - init_time) %>% 
  mutate(part = as.numeric(case_when(
    minus_time <= 90000 ~ 90,
    minus_time > 90000 & minus_time < 120000 ~ 120,
    minus_time >= 120000 ~ 180))) %>%
  mutate(correct_new = ifelse(correct == 0, -1, 1))
```

```{r}
# checking sample size
df.block.time %>%
  select(pid) %>%
  unique() %>%
  tally()
```

## Creating reference dataframe to set intervals for block timing 
```{r}
list <- seq(0, 180000, by = 10000)

intervals.ref <- data.frame(tibble(interval = list))

full.int.df <- df.block.time %>%
  dplyr::rename(seconds_elapsed = minus_time) %>%
  group_by(pid, runId, blockId, seconds_elapsed) %>%
  tally() %>% 
  select(-n) %>%
  merge(intervals.ref, all = TRUE)
```

## Augmenting data for SRE form into 10 Second Intervals
```{r}
df.timing.int <- df.block.time %>%
  select(pid, runId, gradeResponse, blockId, correct, correct_new, minus_time) %>%
  filter(blockId == "lab") %>%
  group_by(pid, runId, blockId) %>%
  dplyr::arrange(minus_time, .by_group = T) %>%
  reframe(seconds_elapsed = minus_time,
         sreScore = cumsum(correct_new))
```

```{r}
df.int.v1 <- df.timing.int %>%
  right_join(full.int.df, by=c("pid", "runId", "blockId", "seconds_elapsed")) %>%
  mutate(diff = interval - seconds_elapsed) %>%
  filter(diff >= 0, blockId == "lab") %>%
  group_by(pid, runId, interval) %>% 
  slice_min(diff) %>%
  dplyr::rename("secondsElapsed" = seconds_elapsed) %>%
  ungroup()
```

```{r}
df.180.score <- df.int.v1 %>%
  filter(interval == 180000) %>%
  select(pid, runId, "sreScore180" = sreScore) %>%
  mutate(source = ifelse(grepl("brs", pid), "BRSSD", "LMB")) 
```

```{r}
df.int.final <- df.int.v1 %>%
  left_join(df.180.score, by=c("pid", "runId")) 
```

```{r}
df.int.final
```

## Cleaning up validation data frames (WJ & TOWRE)

### Removing unused columns from validation assessment dataframes
```{r}
df.lmb.wj.cleaned <- df.lmb.wj.raw %>%
  select(-redcapId) %>%
  mutate(lwidRaw = NA, waRaw = NA, brs = NA,
         sweRaw = NA, pdeRaw = NA)
```

```{r}
df.brs.wj.cleaned <- df.brs.wj.raw %>%
  select(-date, -months, -ends_with("Perc"), -ends_with("SS")) %>%
  mutate(mffScore = NA)
```

### Merging the WJ/TOWRE data for LMB and BRSSD participants
```{r}
df.combined.wj.scores <- rbind(df.lmb.wj.cleaned, df.brs.wj.cleaned) %>%
  mutate(srfScore = srfCorrect - srfIncorrect) %>%
  select(-srfCorrect, -srfIncorrect)
```

### Combining SRE and WJ/TOWRE scores for final analysis dataframe
```{r}
df.sre.validation.analysis <- df.int.final %>%
  left_join(df.combined.wj.scores, by="pid") %>%
  filter(!is.na(srfScore))
```

## Combining SRE & WJ SRF data frame for interval analysis
```{r}
df.sre.srf.int <- df.int.final %>%
  select(pid, runId, sreScore, sreScore180,
         secondsElapsed, interval) %>%
  left_join(df.combined.wj.scores, by = "pid") %>%
  filter(!is.na(srfScore)) %>%
  dplyr::group_by(interval) %>%
  reframe(oneThreeCorr = cor(srfScore, sreScore),
          n = n(), 
          standardError = sqrt((1 - oneThreeCorr^2) / (n - 2)),
          errorBarsMax = oneThreeCorr + standardError,
          errorBarsMin = oneThreeCorr - standardError) %>% 
  mutate(errorBarsMaxAdj = ifelse(errorBarsMax > 1.00, 1.00, errorBarsMax))
```

### Creating data frames for correlations plots (both 1-1 and heat matrix)
```{r}
df.corr.plot <- df.sre.validation.analysis %>%
  select(pid, "age" = years, grade, 
         "sreScore" = sreScore180, lwidRaw, waRaw, srfScore,
         brs, sweRaw, pdeRaw, source) %>% 
  unique()
```

```{r}
df.corr.matrix <- df.corr.plot %>%
  select(-c(pid, age, grade, source)) 

# compute correlations
cor.matrix <- stats::cor(df.corr.matrix, use = "pairwise.complete.obs")

# melt the correlation matrix
df.melted.corr.matrix <- melt(get_upper_tri(cor.matrix), na.rm = TRUE)
```

### Dataframe with mean and standard deviation for all the assessments
```{r}
df.corr.matrix %>%
  dplyr::summarise(sreMean = mean(sreScore),
                   sreSD = sd(sreScore),
                   srfMean = mean(srfScore),
                   srfSD = sd(srfScore),
                   lwidMean = mean(lwidRaw, na.rm = TRUE),
                   lwidSD = sd(lwidRaw, na.rm = TRUE),
                   waMean = mean(waRaw, na.rm = TRUE),
                   waSD = sd(waRaw, na.rm = TRUE),
                   sweMean = mean(sweRaw, na.rm = TRUE),
                   sweSD = sd(sweRaw, na.rm = TRUE),
                   pdeMean = mean(pdeRaw, na.rm = TRUE),
                   pdeSD = sd(pdeRaw, na.rm = TRUE))
```

### Making a dataframe with the sample size for the plots
```{r}
sre.wj.corr.labels <- df.corr.plot %>%
  select(pid, source) %>%
  group_by(source) %>%
  tally() 

df.corr.plot <- df.corr.plot %>%
  left_join(sre.wj.corr.labels, by="source")
```

# Data Visualizations

## Figure 6: Inspecting SRE vs. WJ SRF correlations 
```{r}
school.sre.wj.corr.plot <- ggplot(df.corr.plot %>%
                                    filter(source == "BRSSD"), mapping = aes(x = sreScore,
                                                                     y = srfScore)) + 
  geom_point(aes(color=grade)) + 
  facet_grid(. ~ "3rd grade public school") +
  labs(x = "SRE-Fixed score",
       y = "WJ SRF score") +
  geom_smooth(method = "lm", color="black") +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..), color="black") +
  geom_text(aes(label=paste('n =', n)), 
            data=filter(sre.wj.corr.labels, source=="BRSSD"),
            x=5, y=68, vjust=0, hjust=0) +
  scale_color_gradientn(colours = c('dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12),
        legend.position = "none",
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"))

lmb.sre.wj.corr.plot <- ggplot(df.corr.plot %>%
                                 filter(source == "LMB"), mapping = aes(x = sreScore,
                     y = srfScore)) + 
  geom_point(aes(color = age)) + 
  facet_grid(. ~ "Dyslexia study") +
  labs(x = "SRE-Fixed score",
       y = "WJ SRF score") +
  geom_smooth(method = "lm", color="black") +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..)) +
  geom_text(aes(label=paste('n =', n)), 
            data=filter(sre.wj.corr.labels, source=="LMB"),
            x=5, y=68, vjust=0, hjust=0) +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12),
        legend.position = "none",
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"))

school.sre.wj.corr.plot
lmb.sre.wj.corr.plot
```

```{r}
df.wj.lwid.plot <- df.corr.plot %>% 
  filter(!is.na(pdeRaw))

ggplot(df.corr.plot, mapping = aes(x = sreScore,
                     y = pdeRaw)) + 
  geom_point() + 
  labs(x = "SRE-Fixed score (correct - incorrect)",
       y = "WJ SRF score (correct - incorrect)") +
  geom_smooth(method = "lm", color="black") +
  stat_cor(cor.coef.name = 'r', color = "black", 
           geom = "label", hjust = -0.3, vjust = 1) +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12),
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"),
        aspect.ratio=1)
```

```{r}
sre.wj.int.plot <- ggplot(df.sre.srf.int %>% 
                            filter(interval != 0), mapping = aes(x = interval,
                     y = oneThreeCorr)) + 
  labs(x = "SRE time elapsed (s)",
       y = "Correlation of SRE & WJ-SRF") +
  geom_errorbar(aes(ymin = errorBarsMin, ymax = errorBarsMaxAdj), width = 0.5) +
  geom_smooth(method = "loess", color="gray60", se=F) +
  geom_point(color="gray21") + 
  scale_x_continuous(limits=c(0,180000), breaks=seq(0,180000,30000), labels=seq(0, 180, by=30)) +
  scale_y_continuous(limits = c(0, 1), breaks=seq(0,1.0, 0.25)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12),
        panel.grid.minor.x = element_blank())

sre.wj.int.plot
```

```{r}
# # create a ggheatmap
ggplot(df.melted.corr.matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color="white") +
  scale_fill_gradientn(name = "Pearson Correlation", colours = wes_palette("Zissou1", 100, type = "continuous")) +
  geom_text(aes(Var1, Var2, label = sprintf("%.2f", value)), color = "black", size = 2.75, family="Avenir") +
  theme_light() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1,  size = 8, hjust = 1, face = "bold")) +
  theme(axis.text.y = element_text(size = 8, face = "bold")) +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.title=element_text(size=9, face = "bold"),
        legend.text=element_text(size=5, face = "bold"),
        legend.direction = "horizontal",
        legend.position = "bottom",
        text=element_text(family="Avenir", size=12)) +
  guides(fill = guide_colorbar(barwidth = 5, barheight = .8,
                             title.position = "top", title.hjust = 0.5)) +
  coord_fixed()
```

### Arranging the plots for ggsave
```{r}
combined.plots <- grid.arrange(school.sre.wj.corr.plot, lmb.sre.wj.corr.plot, sre.wj.int.plot, nrow=1)
```

## Saving out plots
```{r}
ggsave('figures/Study 4 Combined Plots.png', combined.plots, dpi=300, height = 3, width = 7.5)
```

