---
title: "Study 3: ROAR-SRE AI form analysis"
output:
  html_document: default
  pdf_document: default
date:  "`r Sys.Date()`"
---
# Loading Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(mgcv)
library(ggplot2)
library(ggpubr)
library(pracma)
library(viridis)
library(wesanderson)
library(lubridate)
```

# Data 

Data was pulled from the ROAR-Research repository on Google Drive. The data can be downloaded [here](https://drive.google.com/drive/folders/1hxPnjGUWaAzcHno1xSDHWK9eWPpUIu3e?usp=drive_link).


## Creating reference dataframes to use later on 
```{r}
list <- seq(0, 180000, by = 10000)

intervals.ref <- data.frame(tibble(interval = list))
```

## Data Loading

```{r SRE-data-load}
# anya's directory 
df.runs <- read.csv("~/Documents/roar-analysis/data/sre-runs-092023.csv")
df.trials <- read.csv("~/Documents/roar-analysis/data/sre-trials-092023.csv")
df.trials.additional <- read.csv("~/Documents/roar-analysis/data/sre-trials-after101723-103123.csv")
df.grade <- read.csv("~/Documents/roar-analysis/data/new-dashboard-user-ages-grades.csv")
```

## Data Wrangling

```{r}
df.invalid.runs.additional <- df.trials.additional %>% 
  filter(completed == "True", variantId == "Kh5eSiySVgLNaGHrFEwu", 
         user.userType == "student", !is.na(blockId), !is.na(user.grade)) %>% 
  group_by(runId, blockId) %>% 
  dplyr::summarise(medianRT = median(rt), 
            pc = mean(correct)) %>%
  filter(pc < 0.65, medianRT < 1000)

df.additional.combined <- df.trials.additional %>% 
  filter(completed == "True",variantId == "Kh5eSiySVgLNaGHrFEwu", user.userType == "student",
         !is.na(blockId), !is.na(user.grade)) %>% 
  filter(!runId %in% df.invalid.runs.additional$runId) %>% 
  group_by(user.assessmentPid, runId, corpusId, user.grade) %>% 
  mutate(correct = ifelse(correct == 1, 1, -1)) %>% 
  dplyr::summarise(totalScore = sum(correct))
```

```{r}
df.invalid.runs <- df.trials %>% 
   filter(completed == "True", user.userType == "student",!is.na(blockId), assessment_stage == "test_response") %>%
   group_by(runId, blockId, user.assessmentPid) %>% 
   dplyr::summarise(medianRT = median(rt), pc = mean(correct)) %>% 
  filter(pc < 0.65, medianRT < 1000) 
```

```{r}
df.runs.selected <- df.runs %>% 
  select(user.assessmentPid, completed, variantId, runId, user.userType,
         scores.computed.lab.sreScore,
         scores.computed.aiV1P1.sreScore, 
         scores.computed.aiV1P2.sreScore)
```

```{r}
df.valid.runs.selected <- df.runs.selected %>% 
  filter(!grepl('-test', user.assessmentPid), variantId == "Kh5eSiySVgLNaGHrFEwu") %>%
  filter(completed == "True", user.userType == "student") %>% 
  left_join(df.grade %>% 
              select(assessmentPid, grade) %>% 
              dplyr::rename(user.assessmentPid = assessmentPid)) %>% 
  filter(grade != "Kindergarten", !is.na(grade)) %>% 
  mutate(grade = as.numeric(grade)) %>% 
  filter(!runId %in% df.invalid.runs$runId) 
```

creating data frame for block timing based off valid trials
```{r}
df.valid.trials <- df.trials %>%
  filter(runId %in% df.valid.runs.selected$runId,
         completed == "True") %>%
  select(user.assessmentPid, runId, corpusId, blockId, itemId, trialNumBlock, trial_index, correct, time_elapsed, rt)
```

```{r}
df.valid.additional.trials <- df.trials.additional %>%
  filter(runId %in% df.additional.combined$runId,
         assessment_stage == "test_response",
         completed == "True") %>%
  select(user.assessmentPid, runId, corpusId, blockId, itemId, trialNumBlock, trial_index, correct, time_elapsed, rt)
```

```{r}
df.valid.interval.trials <- df.valid.trials %>%
  rbind(df.valid.additional.trials)
```

finding the initial start time of each block
```{r}
df.init.time <- df.valid.interval.trials %>% 
  filter(trialNumBlock == 1) %>% 
  mutate(init_time = time_elapsed -  rt) %>%
  select(user.assessmentPid, runId, corpusId, init_time, -time_elapsed, -rt)
```

```{r}
df.block.time <- df.valid.interval.trials %>% 
  select(user.assessmentPid, runId, corpusId, blockId, itemId, correct, time_elapsed, rt) %>% 
  left_join(df.init.time, by = c("user.assessmentPid", "runId", "corpusId")) %>%
  filter(!is.na(user.assessmentPid), user.assessmentPid != "", corpusId != "practice") %>%
  mutate(minus_time = time_elapsed - init_time,
         correct_new = ifelse(correct == 0, -1, 1))
```

```{r}
full.int.df <- df.block.time %>%
  dplyr::rename(seconds_elapsed = minus_time) %>%
  group_by(user.assessmentPid, runId, corpusId, seconds_elapsed) %>%
  tally() %>% 
  select(-n) %>%
  merge(intervals.ref, all = TRUE)
```

```{r}
df.int.select <- df.block.time %>%
  select(user.assessmentPid, runId, corpusId, correct, correct_new, minus_time) %>%
  group_by(user.assessmentPid, runId, corpusId) %>%
  dplyr::arrange(minus_time, .by_group = T) %>%
  reframe(seconds_elapsed = minus_time,
         sreScore = cumsum(correct_new))
```

```{r}
df.int.final <- df.int.select %>%
  right_join(full.int.df, by=c("user.assessmentPid", "runId", "corpusId", "seconds_elapsed")) %>%
  mutate(diff = abs(interval - seconds_elapsed)) %>%
  filter(diff >= 0) %>%
  group_by(user.assessmentPid, runId, corpusId, interval) %>%
  slice_min(diff) %>%
  distinct(runId, corpusId, interval, diff, .keep_all = TRUE) %>%
  dplyr::rename("secondsElapsed" = seconds_elapsed) %>%
  filter(!is.na(sreScore)) %>%
  ungroup()
```

selecting the full score to correlate against intervals
```{r}
df.aiP1.score <- df.int.final %>%
  filter(interval == 180000, corpusId == "aiV1P1") %>%
  select(user.assessmentPid, runId, "aiP1Score180" = sreScore)
```

```{r}
df.aiP2.score <- df.int.final %>%
  filter(interval == 180000, corpusId == "aiV1P2") %>%
  select(user.assessmentPid, runId, "aiP2Score180" = sreScore)
```

```{r}
df.lab.score <- df.int.final %>%
  filter(interval == 180000, corpusId == "lab") %>%
  select(user.assessmentPid, runId, "labScore180" = sreScore)
```

combining ai & lab intervals for timing analysis
```{r}
df.lab.aiP1.int <- df.int.final %>% 
  filter(corpusId == "aiV1P1") %>%
  merge(df.aiP1.score, by=c("user.assessmentPid", "runId")) %>%
  merge(df.lab.score, by=c("user.assessmentPid", "runId")) %>%
  group_by(interval) %>%
  reframe(equalCorr = cor(labScore180, aiP1Score180),
          oneThreeCorr = cor(labScore180, sreScore),
          n = n(), 
          standardError = sqrt((1 - oneThreeCorr^2) / (n - 2)),
          errorBarsMax = oneThreeCorr + standardError,
          errorBarsMin = oneThreeCorr - standardError)
```

```{r}
df.lab.aiP2.int <- df.int.final %>% 
  filter(corpusId == "aiV1P2") %>%
  merge(df.aiP2.score, by=c("user.assessmentPid", "runId")) %>%
  merge(df.lab.score, by=c("user.assessmentPid", "runId")) %>%
  group_by(interval) %>%
  reframe(equalCorr = cor(labScore180, aiP2Score180),
          oneThreeCorr = cor(labScore180, sreScore),
          n = n(), 
          standardError = sqrt((1 - oneThreeCorr^2) / (n - 2)),
          errorBarsMax = oneThreeCorr + standardError,
          errorBarsMin = oneThreeCorr - standardError)
```

# Visualization
```{r}
df.valid.runs.selected.p1 <- df.valid.runs.selected %>% 
  filter(!is.na(scores.computed.aiV1P1.sreScore)) 

df.valid.runs.selected.p2 <- df.valid.runs.selected %>% 
  filter(!is.na(scores.computed.aiV1P2.sreScore)) 
```

```{r}
df.valid.runs.selected.combined <- df.valid.runs.selected.p1 %>% 
  mutate(ai_score = scores.computed.aiV1P1.sreScore, form = "AI-Form A") %>% 
  rbind(df.valid.runs.selected.p2 %>% 
  mutate(ai_score = scores.computed.aiV1P2.sreScore, form = "AI-Form B"))
```

```{r}
t.test( df.valid.runs.selected.combined$ai_score,  df.valid.runs.selected.combined$scores.computed.lab.sreScore)
```

```{r}
df.valid.runs.selected.combined.updated <- df.additional.combined %>% 
  pivot_wider(names_from = corpusId, values_from = totalScore) %>% 
  dplyr::rename(grade = user.grade, scores.computed.lab.sreScore = lab) %>% 
  mutate(ai_score = ifelse(!is.na(aiV1P1), aiV1P1, aiV1P2)) %>% 
  mutate(form = ifelse(!is.na(aiV1P1),"AI-Form A", "AI-Form B")) %>% 
  select(user.assessmentPid, runId,  grade,scores.computed.lab.sreScore, ai_score, form) %>% 
  rbind(df.valid.runs.selected.combined %>% 
  select(user.assessmentPid, runId,  grade,scores.computed.lab.sreScore, ai_score, form))
```

grade distribution
```{r}
df.valid.runs.selected.combined.updated %>% 
  group_by(grade) %>% 
  tally()
```

```{r}
df.n.values <- df.valid.runs.selected.combined.updated %>% 
  group_by(form) %>% 
  dplyr::summarise(n= n(), mean(ai_score), mean(scores.computed.lab.sreScore), 
                   sd(ai_score), sd(scores.computed.lab.sreScore))

df.n.values
```

```{r}
g.sre.ai.compare <- ggplot(df.valid.runs.selected.combined.updated, mapping = aes(y = ai_score,
                     x = scores.computed.lab.sreScore)) + 
  facet_wrap(~form) +
  xlim(-30, 130) + 
  ylim(-30, 130) +
  labs(x = "SRE-Fixed score", 
       y = "SRE-AI score", 
       color = "Grade") + 
  coord_equal() +
  geom_abline(slope=1,intercept = 0) +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..)) +
  geom_text(aes(label=paste('n =', n)), x=-30, y=110, vjust=0, hjust=0, data=df.n.values) +
  geom_point(aes(color=grade), alpha = 0.7, size = 1) + 
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"),
        legend.position = c(0.93, 0.3),
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"))

g.sre.ai.compare
```

```{r}
lab.aiP1.int.plot <- ggplot(df.lab.aiP1.int %>% 
                              filter(interval != 0), 
                            mapping = aes(x = interval, y = oneThreeCorr)) + 
  facet_grid(. ~ "AI-Form A") +
  labs(x = "SRE-AI time elapsed (s)",
       y = "Correlation of SRE-AI and SRE-Fixed") +
  geom_text(aes(family=c("Avenir")), label="n = 458", x=135000, y=0.1) +
  geom_errorbar(aes(ymin = errorBarsMin, ymax = errorBarsMax), width = 0.5) +
  geom_smooth(method = "loess", color="gray60", se=F) +
  geom_point(color="gray21") +
  scale_x_continuous(limits=c(0,180000), breaks=seq(0,180000,60000), labels=seq(0, 180, by=60)) +
  scale_y_continuous(limits = c(0, 1), breaks=seq(0,1.0,0.2)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=14), 
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"))

lab.aiP2.int.plot <- ggplot(df.lab.aiP2.int  %>% 
                              filter(interval != 0),
                            mapping = aes(x = interval, y = oneThreeCorr)) + 
  facet_grid(. ~ "AI-Form B") +
  labs(x = "SRE-AI time elapsed (s)",
       y = "Correlation of SRE-AI and SRE-Fixed") +
  geom_text(aes(family=c("Avenir")), label="n = 436", x=135000, y=0.1) +
  geom_errorbar(aes(ymin = errorBarsMin, ymax = errorBarsMax), width = 0.5) +
  geom_smooth(method = "loess", color="gray60", se=F) +
  geom_point(color="gray21") +
  scale_x_continuous(limits=c(0,180000), breaks=seq(0,180000,60000), labels=seq(0, 180, by=60)) +
  scale_y_continuous(limits = c(0, 1), breaks=seq(0,1.0,0.2)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=14), 
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"))

lab.aiP1.int.plot
lab.aiP2.int.plot

lab.ai.int.combined <- (lab.aiP1.int.plot + lab.aiP2.int.plot)
```

```{r}
g.sre.ai.compare
lab.ai.int.combined
```

```{r}
#ggsave('../figures/SRE-AI.pdf', g.sre.ai.compare)
ggsave('../figures/SRE-AI.png', g.sre.ai.compare, dpi=300, height = 4, width = 7)
#ggsave('../figures/SRE-AI-Interval.pdf', lab.ai.int.combined)
ggsave('../figures/SRE-AI-Interval.png', lab.ai.int.combined, dpi=300, height = 4, width = 7)
```

## grade level 
```{r}
ggplot(df.valid.runs.selected.combined.updated, mapping = aes(y = ai_score,
                     x = scores.computed.lab.sreScore)) + 
  facet_wrap(~grade) +
  geom_point(alpha=0.5, size = 0.2) +
  geom_abline(slope=1,intercept = 0) +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..)) +
  geom_point(aes(color=grade)) + ggtitle('SRE vs. AI Form') +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1'))
```