---
title: "Study 2: ROAR-SRE-TOSREC GAM Analysis"
output:
  html_document: default
  pdf_document: default
date: "2023-05-22"
---
# Loading Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(patchwork)
library(tidyr)
library(dplyr)
library(stringr)
library(mgcv)
library(ggplot2)
library(ggExtra)
library(ggpubr)
library(pracma)
library(equate)
library(viridis)
library(wesanderson)
library(gratia)
library(lubridate)
library(tidygam)
library(patchwork)
library(diptest)
```

# Data 

Data was pulled from the ROAR-Research repository on Google Drive. The data can be downloaded [here](https://drive.google.com/drive/folders/1hxPnjGUWaAzcHno1xSDHWK9eWPpUIu3e?usp=drive_link).

## Data Loading
```{r SRE-data-load}
df.raw <- read.csv('~/Documents/roar-analysis/data/sre-paper-data/updated-sre-dataset-with-rt-pcorr_10-30-2023.csv') # updated dataset for filtering participants
```

```{r}
dim(df.raw)
df.raw
```

## Data Cleaning and Wrangling

### GAM Dataframe Prep 
```{r SRE-data-prep}
# filtering out participants who are missing grade information or the necessary scores
df.cleaned <- df.raw %>%
  filter(!is.na(grade), 
         !is.na(lab.raw.score), 
         !is.na(tosrec.raw.score),
         !is.na(median.rt.lab),
         !is.na(median.rt.tosrec),
         grade <=8) %>% # ONLY keeping participants in grades 1-8 
  mutate(pCorr.overall = (correct.lab + correct.tosrec)/(attempted.lab + attempted.tosrec),
         attempted.overall = attempted.lab + attempted.tosrec)
```

```{r SRE-data-prep}
# filtering out participants who are missing grade information or the necessary scores
df.cleaned.include.upper <- df.raw %>%
  filter(!is.na(grade), 
         !is.na(lab.raw.score), 
         !is.na(tosrec.raw.score),
         grade > 8) %>% # keeping participants in grades 1-12 
  mutate(pCorr.overall = (correct.lab + correct.tosrec)/(attempted.lab + attempted.tosrec),
         attempted.overall = attempted.lab + attempted.tosrec)
```

```{r SRE-data-prep}
dim(df.cleaned)
df.cleaned 
```

### Filtering out participants before analysis
```{r}
# filtering out guessing participants who have low accuracy and fast reaction times
df.filtered <- df.cleaned %>%
  filter(((median.rt.lab > 1000 & median.rt.lab < 20000) & pCorr.lab > 0.65) | 
  ((median.rt.tosrec > 1000 & median.rt.tosrec < 20000) & pCorr.tosrec > 0.65)) %>%
  mutate(pidPrefix = str_extract(pid, "[^-]+")) %>%
  mutate(group = ifelse(grepl("brs|wsd|kex", pid), "Public", "Dyslexia"))

df.filtered.upper <- df.cleaned.include.upper %>%
  filter(((median.rt.lab > 1000 & median.rt.lab < 20000) & pCorr.lab > 0.65) | 
  ((median.rt.tosrec > 1000 & median.rt.tosrec < 20000) & pCorr.tosrec > 0.65))
```

# Analysis 

### Checking median pCorr and interquartile ranges RT for SRE-Fixed and TOSREC
```{r}
# median proportion correct 
paste0("sre-fixed median proportion correct:", median(df.cleaned$pCorr.lab))
paste0("tosrec median proportion correct:", median(df.cleaned$pCorr.tosrec))

# interquartile ranges for median RT
quantile(df.cleaned$median.rt.lab)
quantile(df.cleaned$median.rt.tosrec)
```

### Checking if proportion correct and reaction time for SRE-Fixed are bimodal
```{r}
dip.test(df.cleaned$median.rt.lab)
dip.test(df.cleaned$pCorr.lab)
```

### Checking if proportion correct and reaction time for TOSREC are bimodal
```{r}
dip.test(df.cleaned$median.rt.tosrec)
dip.test(df.cleaned$pCorr.tosrec)
```

Figure 4: Inspecting the distribution of proportion correct vs. reaction time for SRE-Fixed and TOSREC 
```{r}
lab.rt.pcorr.plot <- ggplot(df.cleaned, aes(x = pCorr.lab, y = median.rt.lab, color=tosrec.grade.response)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept=1000, linetype="solid", 
             color = "red", size=0.5) +
  geom_vline(xintercept=0.65, linetype="solid", 
             color = "red", size=0.5) +
  labs(x = "SRE-Fixed proportion correct",
       y = "Median RT (milliseconds)",
       color = "Grade") + 
  scale_y_continuous(limits = c(0, 20000), breaks = seq(0, 20000, 5000)) +
  scale_x_continuous(limits = c(0.3, 1.0), breaks = seq(0.3, 1.0, 0.1)) +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=14), 
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"), 
        legend.position = "none", 
        aspect.ratio=1) 

tosrec.rt.pcorr.plot <- ggplot(df.cleaned, aes(x = pCorr.tosrec, y = median.rt.tosrec, color=tosrec.grade.response)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept=1000, linetype="solid", 
             color = "red", size=0.5) +
  geom_vline(xintercept=0.65, linetype="solid", 
             color = "red", size=0.5) +
  labs(x = "TOSREC proportion correct",
       y = "Median RT (milliseconds)",
       color = "Grade") + 
  scale_y_continuous(limits = c(0, 20000), breaks = seq(0, 20000, 5000)) +
  #scale_y_continuous(limits = c(0, 15000), breaks = seq(0, 15000, 2000)) +
  scale_x_continuous(limits = c(0.3, 1.0), breaks = seq(0.3, 1.0, 0.1)) +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=14), 
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold"), 
        legend.position = "none", 
        aspect.ratio=1) 

lab.rt.pcorr.marg.plot <- ggMarginal(lab.rt.pcorr.plot, type="histogram")
tosrec.rt.pcorr.marg.plot <- ggMarginal(tosrec.rt.pcorr.plot, type="histogram")

# viewing the plots
lab.rt.pcorr.marg.plot
tosrec.rt.pcorr.marg.plot
```

Figure: Inspecting SRE-Fixed score vs. TOSREC score (Public vs. Dyslexia)
```{r SRE-GAM-LOESS-Plot}
school.compare.plot <- ggplot(df.filtered, aes(y = lab.raw.score, x = tosrec.raw.score, color=tosrec.grade.response)) +
  facet_wrap(~group) +
  geom_point(alpha=0.50) +
  geom_abline(slope=1,intercept = 0) +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..)) +
  labs(x = "ROAR-SRE score (correct - incorrect)",
       y = "TOSREC score (correct - incorrect)") +
  theme_light() +
  theme(text=element_text(family="Avenir", size=14), 
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"),
        aspect.ratio=1)

school.compare.plot
```

Figure: Inspecting SRE-Fixed score vs. TOSREC score for grades 9-12
```{r SRE-GAM-LOESS-Plot}
# Plot GAM prediction
upper.cor.plot <- ggplot(df.filtered.upper, aes(y = lab.raw.score, x = tosrec.raw.score, color=tosrec.grade.response)) +
  geom_point(alpha=0.50) +
  geom_abline(slope=1,intercept = 0) +
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..)) +
  labs(x = "ROAR-SRE score (correct - incorrect)",
       y = "TOSREC score (correct - incorrect)") +
  theme_light() +
  theme(text=element_text(family="Avenir", size=14), 
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"),
        aspect.ratio=1)

upper.cor.plot
```

Figure: Inspecting meta-data grade (reported by school partners) against the grade response for SRE
```{r SRE-data-prep}
df.grade.check <- df.filtered

ggplot(df.grade.check, aes(x=grade,y=tosrec.grade.response)) +
  geom_bin_2d() + 
  scale_fill_viridis(option ='magma',direction=-1) +
  scale_x_continuous(breaks = c(1:12)) +
  scale_y_continuous(breaks = c(1:12))
```

Figure: Grade distribution plot of Study 2 ROAR-SRE participant sample 
```{r SRE-data-prep}
df.grade.dist <- df.filtered

ggplot(df.grade.dist, aes(x=tosrec.grade.response)) +
  geom_bar(position="identity", fill="gray55", col=I("black"), width=1) + 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, color = "black", size = 4, family = "Avenir") +
  labs(x = "Grade",
       y = "Count of participants") +
  scale_y_continuous(limits = c(0, 450), expand = c(0,0)) +
  scale_x_continuous(n.breaks = 8, expand = c(0, 0)) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12)) 
```
### Demographics distribution of schools ###
```{r}
df.school.demo <- df.filtered %>%
  mutate(schoolPrefix = str_extract(pid, "^[^-]+")) %>%
  select(schoolPrefix) %>%
  unique()
```

### Assigning GAM model data to final variable for easy de-bugging
```{r}
df.final <- df.filtered
```

## Fitting GAM & local regression model (Loess) to SRE data
```{r SRE-Fit-GAM-LOESS}
# gr.gam <- gam(equivalent.tosrec.ss ~ te(lab.raw.score, tosrec.grade.response, k=3), bs='tp', data=df.final, method='REML')
# we're using the default parameters for the paper, but keeping the k = 3 for the lookup table
gr.gam <- gam(equivalent.tosrec.ss ~ te(lab.raw.score, tosrec.grade.response), data=df.final, method='REML')
gr.loess <- loess(equivalent.tosrec.ss ~ lab.raw.score*tosrec.grade.response, data=df.final)
```

### Checking summary of GAM & Loess
```{r SRE-GAM-LOESS-Summary}
summary(gr.gam)
summary(gr.loess)
```

## Predicting TOSREC standard score for each data point
```{r SRE-Fit-GAM-LOESS}
df.final$predicted.tosrec.gam <- predict(gr.gam, df.final)
df.final$predicted.tosrec.loess <- predict(gr.loess, df.final)
```

```{r SRE-Fit-GAM-LOESS}
sprintf('GAM prediciton is correlated with TOSREC SS at r= %f', cor(df.final$predicted.tosrec.gam, df.final$equivalent.tosrec.ss, use="pairwise.complete.obs"))
sprintf('LOESS prediciton is correlated with TOSREC SS at r= %f', cor(df.final$predicted.tosrec.loess, df.final$equivalent.tosrec.ss, use="pairwise.complete.obs"))
```

```{r}
# creating sample size labels for each grade to use in the plot below
grade.labels <- df.final %>% 
  group_by(tosrec.grade.response) %>% 
  tally()
```

Figure 5: Inspecting predicted TOSREC standard score vs. equivalent TOSREC standard score
```{r SRE-GAM-LOESS-Plot}
# Plot GAM prediction
gam.pred.plot = ggplot(df.final, aes(y = predicted.tosrec.gam, x = equivalent.tosrec.ss, color=tosrec.grade.response)) +
  geom_abline(slope=1,intercept = 0,alpha = 0.5) +
  geom_point(alpha=0.50) +
  facet_wrap(~tosrec.grade.response, nrow=2, labeller = labeller(tosrec.grade.response = c("1"="1st grade", "2"="2nd grade", "3"="3rd grade", "4"="4th grade", "5"="5th grade", "6"="6th grade", "7"="7th grade", "8"="8th grade"))) + 
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..)) +
  geom_text(aes(label=paste('n =', n)), x=50, y=125, vjust=0, hjust=0, data=grade.labels) +
  labs(x = "ROAR-SRE Standard Score",
       y = "TOSREC Standard Score") + 
  xlim(50,145) + 
  ylim(50,145) +  
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=14), 
        legend.position = "none",
        axis.title = element_text(face = "bold"),
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"),
        aspect.ratio=1)

# Plot loess prediction
loess.pred.plot = ggplot(df.final, aes(predicted.tosrec.loess, equivalent.tosrec.ss, color=tosrec.grade.response)) +
  geom_point(alpha=0.5) +
  geom_abline(slope=1,intercept = 0) +
  facet_wrap(~tosrec.grade.response, nrow=2, labeller = labeller(tosrec.grade.response = c("1"="1st grade", "2"="2nd grade", "3"="3rd grade", "4"="4th grade", "5"="5th grade", "6"="6th grade", "7"="7th grade", "8"="8th grade"))) + 
  stat_cor(cor.coef.name = 'r', aes(label = ..r.label..)) +
  labs(x = "ROAR-SRE Standard Score",
       y = "TOSREC Standard Score") + 
  xlim(50,145) + 
  ylim(50,145) + 
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1')) +
  theme_light() +
  theme(text=element_text(family="Avenir", size=12), 
        legend.position = "none",
        strip.text = element_text(colour = 'black'),
        strip.background = element_rect(fill="gray85"),
        aspect.ratio=1)

# Viewing the plots
gam.pred.plot
loess.pred.plot
```

Figure: Inspecting the density of TOSREC scores by grades
```{r SRE-TOSRECT-Lookup-Table}
# Density of TOSREC scores by grade
gam.ss.dens.plot <- ggplot(df.final, aes(x=equivalent.tosrec.ss,
                                         group=tosrec.grade.response)) +
  geom_density(aes(color=tosrec.grade.response)) +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1'))

gam.raw.dens.plot <- ggplot(df.final, aes(x=lab.raw.score, group=tosrec.grade.response)) +
  geom_density(aes(color=tosrec.grade.response))+
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1'))

gam.ss.dens.plot
gam.raw.dens.plot
```

## Evaluating the GAM and Loess for each TOSREC & grade combination
```{r SRE-TOSRECT-Lookup-Table}
grid <- meshgrid(x=1:8, y=0:120)
X<- tibble(tosrec.grade.response = as.vector(grid$X))
X$lab.raw.score <- as.vector(grid$Y)
X$equivalent.tosrec.ss.gam <- predict(gr.gam, X)
X$equivalent.tosrec.ss.loess <- predict(gr.loess, X)
```

Figure: Inspecting the GAM and Loess equate
```{r SRE-TOSRECT-Lookup-Table}
# Make a plot of the GAM and the loess equate
ggplot(X, aes(x=lab.raw.score, y=equivalent.tosrec.ss.gam, group=tosrec.grade.response))+
  geom_line(aes(color=tosrec.grade.response)) + xlim(0,125) + ylim(50,145) + ggtitle('GAM Prediction') +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1'))

ggplot(X, aes(x=lab.raw.score, y=equivalent.tosrec.ss.loess, group=tosrec.grade.response))+
  geom_line(aes(color=tosrec.grade.response)) + xlim(0,125) + ylim(50,145) + ggtitle('LOESS Prediction') +
  scale_color_gradientn(colours = c( 'dodgerblue1','firebrick1','goldenrod1'))
```

# Run this chunk to save the plots 
```{r}
ggsave('figures/SRE-TOSREC (GAM).png', gam.pred.plot, dpi=300)
ggsave('figures/SRE-TOSREC (Loess).png', loess.pred.plot, dpi=300)
ggsave('figures/SRE-pCorr-RT.png', lab.rt.pcorr.marg.plot, dpi=300)
ggsave('figures/TOSREC-pCorr-RT.png', tosrec.rt.pcorr.marg.plot, dpi=300)
```

